---
title: "Week2-2"
author: "Kyu Cho"
date: "Tuesday, June 16, 2015"
output: html_document
---
# Moneyball
```{r cache=TRUE}
setwd("D:/Google Drive/College/4-The Analytics Edge/data2")
baseball <- read.csv("baseball.csv")
str(baseball)
```

```{r cache=TRUE}
moneyball = subset(baseball, Year < 2002)
str(moneyball)
```

Now confirm the claim made in Moneyball that a team needs to score at least 135 more runs than they allow to win at least 95 games.

```{r cache=TRUE}
plot(moneyball$W, moneyball$Playoffs)
t = table(moneyball$W > 95, moneyball$Playoffs)
# Accuracy
(t[1,1]+t[2,2])/sum(t)

# Build a linear regression equation to predict wins using the difference between runs scored and runs allowed.
moneyball$RD = moneyball$RS - moneyball$RA
str(moneyball)

plot(moneyball$RD, moneyball$W)
WinsReg = lm(W ~ RD, data = moneyball)
summary(WinsReg)
```
You can see the strong correlation between RD and W up to .88

We get the linear regression model.  
Wins = Intercept term + coefficients RD * RD  
W = 80.8814 + 0.1058(RD)  
W >= 95  
80.8814 + 0.1058(RD) >= 95  
RD >= (95-80.8814)/.1058  = 133.4  

This tells us that if the run difference of a team is greater than or equal to 133.4, then we predict that the team will win at least 95 games.  


Q: If a baseball team scores 713 runs and allows 614 runs, how many games do we expect the team to win?  

```{r}
RD2 = 713-614
W2 = 80.8814 + .1058*RD2
W2
```

Now
1. Find how many runs a team will score by using batting statistics.
2. Fine how many runs a team will allow by using fielding and pitching statistic.

 - OBP : On-Base percentage
 - SLG : Slugging percentage
 - BA : Batting Avg.
 
## Scoring Runs by using batting statistics 
```{r cache=TRUE}
RunsReg = lm(RS ~ OBP + SLG + BA, data = moneyball)
summary(RunsReg)
```

As you see, the coefficients for Batting average is actually negative, which implies that a team with a lower batting average will score more runs, which is non-sense. Let's investigate more indepth by removing BA variable.
```{r cache=TRUE}
RunsReg2 = lm(RS ~ OBP + SLG, data = moneyball)
summary(RunsReg2)
```
It looks no difference between the previous model, in fact, it looks more simplier

Thus, we confirmed that Batting Average is overvalued and On-Base percentage is the most important and Slugging Percentage is the second most important by using Linear Regression model.

We get the linear regression model.  
Runs Scored = Intercept term + coeff. OBP * (OBP) + coeff. SLG * (SLG)  
RS = -804.63 + 2737.77(OBP) + 1584.91(SLG)  

**Q: If a baseball team's OBP is 0.311 and SLG is 0.405, how many runs do we expect the team to score?**
```{r cache=TRUE}
RS = -804.63 + 2737.77*.311 + 1584.91*.405
RS
```


## Allowing Runs by using pitching statistic
 - OOBP : Opponents On-Base Percentage
 - OSLG : Opponents Slugging Percentage
 
```{r cache=TRUE}
PitchReg = lm(RA ~ OOBP + OSLG, data = moneyball)
summary(PitchReg)
```
We get the linear regression model.   
Runs Allowed = Intercept term + coeff. OOBP * OOBP + coeff. OSLG * OSLG  
RA = -837.38 + 2913.60(OOBP) + 1514.29(OSLG)  

R^2 = .91 and both variables are significants

**Q: If a baseball team's opponents OBP (OOBP) is 0.297 and oppenents SLG (OSLG) is 0.370, how many runs do we expect the team to allow?**
```{r cache=TRUE}
RA = -837.38 + 2913.60*0.297 + 1514.29*0.370
RA
```
 
Now use our previous formula to predict the wins  
Wins = Intercept term + coefficients RD * RD  
W = 80.8814 + 0.1058(RD)  
RD = (RS - RA)  
```{r cache=TRUE}
RD = RS - RA
W = 80.8814 + 0.1058 * RD
W
```
It predicts that we will win 92 games.

**Problem**  
Need a team that make it to the play off.  
**Objective**  
Build the model that calculates whether or not the win will make it to the play off.  
**Actual Method Used**  
Linear regression model.  
**Procedure**  
1.	Fine the number of win to make to the play off  
-	Use the general statistic method to find, (95 wins is required with accuracy of 90% of chance)  
2.	Build the model that calculates the RD(RS - RA) values to make 95 wins.  
3.	Build the model that calculates the RS based on the player's stats. (OBP, SLG)  
4.	Build the model that calculates the RA based on the opponent's stats (OOBP, OSLG)  
**Facts Discovered**  
95 games is needed to win.  
133 run difference to make 95 games.  
OBP, SLG variables helps to found the run difference.  
