---
title: "Week2-3"
author: "Kyu Cho"
date: "Wednesday, June 17, 2015"
output: html_document
---

```{r cache=TRUE}
setwd("D:/Google Drive/College/4-The Analytics Edge/data2")
NBA = read.csv("NBA_train.csv")
NBA_test = read.csv("NBA_test.csv")
str(NBA)
```

```{r cache=TRUE}
tail(table(NBA$W, NBA$Playoffs),40)
```
Looks like wining 42 games would make it the playoff.  

Let's find out the variable that has strong linear relationship, start with point difference.  
```{r cache=TRUE}
NBA$PTSdiff = NBA$PTS - NBA$oppPTS
plot(NBA$PTSdiff, NBA$W)
```

Since, it looks they have strong linear relationship, linear regression is the good way to start.  

```{r cache=TRUE}
WinsReg = lm(W ~ PTSdiff, data = NBA)
summary(WinsReg)
```
As we assumed, it has R-squared of .9423 so it's very strong relationship.  

Let's make linear equation out of it.  
W = 41 + 0.0326 * PTSdiff  
We need at least 42 wins in order to make it to the playoff safely.
W >= 42
PTSdiff >= (42-41)/0.0326 = 30.63
We need to score at least 31 more in order to make to the playoff.

```{r cache=TRUE}
PointsReg = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + DRB + TOV + STL + BLK, data = NBA)
summary(PointsReg)

head(PointsReg$residuals)
SSE = sum(PointsReg$residual^2)
SSE

# sum of sqare error number(SSE) isn't really interpretable so let's find root mean square erro(RMSE), which is more interpretable.
RMSE = sqrt(SSE/nrow(NBA))
RMSE
mean(NBA$PTS)
```
So making error of 184 points out of average points of 8370 isn't that bad.  

Let's remove some of insignificant variables from the LM.  
Let's remove the variable that has largest Pr(>|t|) value.
```{r}
# removing TOV
PointsReg2 = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + DRB + STL + BLK, data = NBA)
summary(PointsReg2)

# removing DRB
PointsReg3 = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + STL + BLK, data = NBA)
summary(PointsReg3)

# removing BLK
PointsReg4 = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + STL, data = NBA)
summary(PointsReg4)
```

```{r cache=TRUE}
SSE_4 = sum(PointsReg4$residuals^2)
SSE_4
RMSE_4 = sqrt(SSE_4/nrow(NBA))
RMSE_4
```

We are able to keep the RMSE about the same yet looks more simplier and more interpretable.  

```{r cache=TRUE}
str(NBA_test)
PointsPredictions = predict(PointsReg4, newdata=NBA_test)  
SSE = sum((PointsPredictions - NBA_test$PTS)^2)  
SST = sum((mean(NBA$PTS) - NBA_test$PTS)^2)  
R2 = 1 - SSE/SST  
RMSE = sqrt(SSE/nrow(NBA_test))  

PointsPredictions
SSE
SST
R2
RMSE
```

