---
title: "Week7-4 Election Forcasting"
author: "Kyu Cho"
date: "July 16, 2015"
output: html_document
---
# Introduction
construct US presidential election predictions. We separated our data into a training set, containing data from 2004 and 2008 polls, and a test set, containing the data from 2012 polls. We then proceeded to develop a logistic regression model to forecast the 2012 US presidential election.  

In this homework problem, we'll revisit our logistic regression model from Unit 3, and learn how to plot the output on a map of the United States. Unlike what we did in the Crime lecture, this time we'll be plotting predictions rather than data!  

```{r cache=T}
library(ggplot2)
library(caTools)
setwd("D:/Google Drive/College/4-The Analytics Edge/data7")
polling = read.csv("PollingImputed.csv")
str(polling)
```

# DRAWING A MAP OF THE US
```{r cache=T}
statesMap = map_data("state")
str(statesMap)

# Group, defines the different shapes or polygons on the map.
# How many different groups are there?
nrow(table(statesMap$group))

ggplot(statesMap, aes(x=long, y=lat, group=group)) +
    geom_polygon(fill="white", color="black")
```

# COLORING THE STATES BY PREDICTIONS  
```{r cache=TRUE}
train = subset(polling, Year >= 2004 & Year <= 2008)
test = subset(polling, Year >= 2012)

model = glm(Republican~SurveyUSA+DiffCount, data=train, family="binomial")
pred = predict(model, newdata=test, type="response")

predBinary = as.numeric(pred > 0.5)
predDF = data.frame(pred, predBinary, test$State)

# For how many states is our binary prediction 1 (for 2012), corresponding to Republican?
sum(predBinary)

# What is the average predicted probability of our model (on the Test set, for 2012)?
mean(predBinary)
```

## Merge Data frame with "stateMap"
```{r cache=T}
# Unify the state name before the combination.
predDF$region = tolower(predDF$test.State)

# Merge, some of observations will be removed, if the variable "region" is not co-exist on both datasets.
predMap = merge(statesMap, predDF, by="region")

# Order
predMap = predMap[order(predMap$order), ]
```

## Coloring
```{r cache=T}
ggplot(predMap, aes(x=long, y=lat, group=group, fill=predBinary)) +
    geom_polygon(color="black")

# See the color options
head(colors())

ggplot(predMap, aes(x=long, y=lat, group=group, fill=predBinary)) + 
    geom_polygon(color="black") + 
    scale_fill_gradient(low="blue", high="red", guide="legend", breaks= c(0,1), labels=c("Democrat", "Republican"), name="Prediction 2012")
```

# UNDERSTANDING THE PREDICTIONS
```{r cache=T}
# What was our predicted probability for the state of Florida?
which(test$State=="Florida")
pred[6]
```

# PARAMETER SETTINGS
```{r cache=TRUE}
ggplot(predMap, aes(x=long, y=lat, group=group, fill=predBinary)) + 
    geom_polygon(color="black", linetype=1, size=2, alpha=0.3) + 
    scale_fill_gradient(low="blue", high="red", guide="legend", breaks= c(0,1), labels=c("Democrat", "Republican"), name="Prediction 2012")
```

