---
title: "Week7-6 Word Clouds"
author: "Kyu Cho"
date: "July 18, 2015"
output: html_document
---
# Introduction
Earlier in the course, we used text analytics as a predictive tool, using word frequencies as independent variables in our models. However, sometimes our goal is to understand commonly occurring topics in text data instead of to predict the value of some dependent variable. In such cases, word clouds can be a visually appealing way to display the most frequent words in a body of text.  

A word cloud arranges the most common words in some text, using size to indicate the frequency of a word. For instance, this is a word cloud for the complete works of Shakespeare, removing English stopwords:  


# Variables
- Tweet : the text of the tweet.  
-  Avg : the sentiment of the tweet, as assigned by users of Amazon Mechanical Turk. The score ranges on a scale from -2 to 2, where 2 means highly positive sentiment, -2 means highly negative sentiment, and 0 means neutral sentiment. 

```{r cache=T}
library(NLP)
library(tm)
library(wordcloud)
library(RColorBrewer)
setwd("D:/Google Drive/College/4-The Analytics Edge/data7")
tweets = read.csv("tweets.csv")
str(tweets)
```

# PREPARING THE DATA
```{r cache=T}
# We are skiping stem process, It will be easier to read and understand the word cloud if it includes full words instead of just the word stems.
corpus = Corpus(VectorSource(tweets$Tweet))
corpus = tm_map(corpus, tolower)
corpus = tm_map(corpus, PlainTextDocument)
corpus = tm_map(corpus, removePunctuation)
corpus = tm_map(corpus, removeWords, stopwords(c("apple", "english")))

# Create Matrix
mTweets = DocumentTermMatrix(corpus)
mTweets

# Convert into data frame
allTweets = as.data.frame(as.matrix(mTweets))
```

# BUILDING A WORD CLOUD
```{r cache=T}
wordcloud(colnames(allTweets), colSums(allTweets))

# Negative tweet
negativeTweets = subset(allTweets, tweets$Avg <= -1)

display.brewer.all() 

# 30% of words are roated
colors=brewer.pal(9, "Blues")[c(-1, -2, -3, -4)]
wordcloud(colnames(negativeTweets), colSums(negativeTweets),min.freq=3, random.order=F, rot.per=.3, colors=colors)
```
