---
title: "Week3-6"
author: "Kyu Cho"
date: "Monday, June 29, 2015"
output: html_document
---
# Introduction
In the lending industry, investors provide loans to borrowers in exchange for the promise of repayment with interest. If the borrower repays the loan, then the lender profits from the interest. However, if the borrower is unable to repay the loan, then the lender loses money. Therefore, lenders face the problem of predicting the risk of a borrower being unable to repay a loan.

To address this problem, we will use publicly available data from LendingClub.com, a website that connects borrowers and investors over the Internet. This dataset represents 9,578 3-year loans that were funded through the LendingClub.com platform between May 2007 and February 2010. The binary dependent variable not_fully_paid indicates that the loan was not paid back in full (the borrower either defaulted or the loan was "charged off," meaning the borrower was deemed unlikely to ever pay it back).

# Variables
- credit.policy: 1 if the customer meets the credit underwriting criteria of LendingClub.com, and 0 otherwise.
- purpose: The purpose of the loan (takes values "credit_card", "debt_consolidation", "educational", "major_purchase", "small_business", and "all_other").
- int.rate: The interest rate of the loan, as a proportion (a rate of 11% would be stored as 0.11). Borrowers judged by LendingClub.com to be more risky are assigned higher interest rates.
- installment: The monthly installments ($) owed by the borrower if the loan is funded.
- log.annual.inc: The natural log of the self-reported annual income of the borrower.
- dti: The debt-to-income ratio of the borrower (amount of debt divided by annual income).
- fico: The FICO credit score of the borrower.
days.with.cr.line: The number of days the borrower has had a credit line.
- revol.bal: The borrower's revolving balance (amount unpaid at the end of the credit card billing cycle).
- revol.util: The borrower's revolving line utilization rate (the amount of the credit line used relative to total credit available).
inq.last.6mths: The borrower's number of inquiries by creditors in the last 6 months.
- delinq.2yrs: The number of times the borrower had been 30+ days past due on a payment in the past 2 years.
- pub.rec: The borrower's number of derogatory public records (bankruptcy filings, tax liens, or judgments).


```{r cache=TRUE}
setwd("D:/Google Drive/College/4-The Analytics Edge/data3")
loans = read.csv("loans.csv")
loans2 = read.csv("loans_imputed.csv")
summary(loans)
str(loans)
head(loans)
```

# PREPARING THE DATASET
```{r cache=TRUE}
# What proportion of the loans in the dataset were not paid in full?
table(loans$not.fully.paid)
1533/(1533+8045)
```

List of missing data.
- log.annual.inc  
- days.with.cr.line  
- revol.util  
- inq.last.6moths  
- delinq.2yrs  
- pub.rec  
We want to be able to predict risk for all borrowers, instead of just the ones with all data reported.  

# Filing missing data
Imputation predicts missing variable values for a given observation using the variable values that are reported. We called the imputation on a data frame with the dependent variable not.fully.paid removed, so we predicted the missing values using only other independent variables.

```{r cache=TRUE}
library(mice)
set.seed(144)

vars.for.imputation = setdiff(names(loans), "not.fully.paid")
imputed = complete(mice(loans[vars.for.imputation]))
loans[vars.for.imputation] = imputed
```

# PREDICTION MODELS
```{r cache=TRUE}
library(caTools)
set.seed(144)

# Spilt data into train and test sets
split = sample.split(loans$not.fully.paid, SplitRatio = 0.7)

train = subset(loans, split==TRUE)
test = subset(loans, split==FALSE)
#or 
train <- loans[split,]
test <- loans[!split,]

# Build model
loansLog = glm(not.fully.paid ~ ., data=train, family="binomial")
summary(loansLog)
```

```{r cache=TRUE}
# Let Logit(A) be the log odds of loan A not being paid back in full, according to our logistic regression model, and define Logit(B) similarly for loan B. What is the value of Logit(A) - Logit(B)?  
log = -9.406e-03 * (700-710)

# Now, let O(A) be the odds of loan A not being paid back in full, according to our logistic regression model, and define O(B) similarly for loan B. What is the value of O(A)/O(B)? 
exp(log)
```
Using the answer from the previous question, the predicted odds of loan A not being paid back in full are exp(0.09317) = 1.0976 times larger than the predicted odds for loan B. Intuitively, it makes sense that loan A should have higher odds of non-payment than loan B, since the borrower has a worse credit score.  

#Find sensitivity, specifity, accuracy, baseline accurary
```{r cache=TRUE}
predicted.risk = predict(loansLog, newdata=test, type="response")
test$predicted.risk = predicted.risk

TP <- sum(predicted.risk > .5 & test$not.fully.paid)
FN <- sum(predicted.risk <= .5 & test$not.fully.paid)
sensitivity <- TP / (TP+FN)
sensitivity

TN <- sum(predicted.risk <= .5 & !test$not.fully.paid)
FP <- sum(predicted.risk > .5 & !test$not.fully.paid)
specifity <- TN / (TN+FP)
specifity

accuracy <- (TP + TN) / nrow(test)
accuracy

accuracy_baseline <- sum(!test$not.fully.paid) / nrow(test)
accuracy_baseline

# Use the ROCR package to compute the test set AUC
library(ROCR)
ROCRpred <- prediction(test$predicted.risk, test$not.fully.paid)
as.numeric(performance(ROCRpred, "auc")@y.values)
```
The model has poor accuracy at the threshold 0.5. But despite the poor accuracy, we will see later how an investor can still leverage this logistic regression model to make profitable investments.  

# A "SMART BASELINE"  
In the previous problem, we built a logistic regression model that has an AUC significantly higher than the AUC of 0.5 that would be obtained by randomly ordering observations.  
However, LendingClub.com assigns the interest rate to a loan based on their estimate of that loan's risk. This variable, int.rate, is an independent variable in our dataset. In this part, we will investigate using the loan's interest rate as a "smart baseline" to order the loans according to risk.  
```{r cache=TRUE}
loansLog2 = glm(not.fully.paid ~ int.rate, data=train, family="binomial")
summary(loansLog2)
```
As we can see, int.rate is highly correlated with othe risk-related variables, and therefore does not incrementally improve the model when those other variables are included at the 0.05 level in the model trained with all the independent variable.  But now it shows that it's highly significant in the bivariate model.  

```{r cache=TRUE}
predicted.risk2 = predict(loansLog2, newdata=test, type="response")
max(predicted.risk2)

ROCRpred <- prediction(predicted.risk2, test$not.fully.paid)
as.numeric(performance(ROCRpred, "auc")@y.values)
```

# COMPUTING THE PROFITABILITY OF AN INVESTMENT
While thus far we have predicted if a loan will be paid back or not, an investor needs to identify loans that are expected to be profitable. If the loan is paid back in full, then the investor makes interest on the loan. However, if the loan is not paid back, the investor loses the money invested. Therefore, the investor should seek loans that best balance this risk and reward.

How much does a $10 investment with an annual interest rate of 6% pay back after 3 years, using continuous compounding of interest? Hint: remember to convert the percentage to a proportion before doing the math. Enter the number of dollars, without the $ sign.

```{r cache=TRUE}
# To compute interest revenue, consider a $c investment in a loan that has an annual interest rate r over a period of t years. Using continuous compounding of interest, this investment pays back c * exp(rt) dollars by the end of the t years, where exp(rt) is e raised to the r*t power.
10*exp(.06*3)

# Caclulate total only the interest money
10*exp(.06*3) - 10
```

# A SIMPLE INVESTMENT STRATEGY
In the previous subproblem, we concluded that an investor who invested c dollars in a loan with interest rate r for t years makes c * (exp(rt) - 1) dollars of profit if the loan is paid back in full and -c dollars of profit if the loan is not paid back in full (pessimistically).  

In order to evaluate the quality of an investment strategy, we need to compute this profit for each loan in the test set. For this variable, we will assume a $1 investment (aka c=1). To create the variable, we first assign to the profit for a fully paid loan, exp(rt)-1, to every observation, and we then replace this value with -1 in the cases where the loan was not paid in full. All the loans in our dataset are 3-year loans, meaning t=3 in our calculations.  

```{r cache=TRUE}
test$profit = exp(test$int.rate*3) - 1
test$profit[test$not.fully.paid == 1] = -1

# What is the maximum profit of a $10 investment in any loan in the testing set
10*max(test$profit)
```

# AN INVESTMENT STRATEGY BASED ON RISK
A simple investment strategy of equally investing in all the loans would yield profit $20.94 for a $100 investment. But this simple investment strategy does not leverage the prediction model we built earlier in this problem. As stated earlier, investors seek loans that balance reward with risk, in that they simultaneously have high interest rates and a low risk of not being paid back.  

To meet this objective, we will analyze an investment strategy in which the investor only purchases loans with a high interest rate (a rate of at least 15%), but amongst these loans selects the ones with the lowest predicted risk of not being paid back in full. We will model an investor who invests $1 in each of the most promising 100 loans.


```{r cache=TRUE}
highInterest = subset(test, int.rate >= 0.15)
# What is the average profit of a $1 investment in one of these high-interest loans?
mean(highInterest$profit)

# What proportion of the high-interest loans were not paid back in full?
table(highInterest$not.fully.paid)
110/(327+110)
```

Next, we will determine the 100th smallest predicted probability of not paying in full by sorting the predicted risks in increasing order and selecting the 100th element of this sorted list.  
```{r cache=TRUE}
cutoff = sort(highInterest$predicted.risk, decreasing=FALSE)[100]
selectedLoans = subset(highInterest, predicted.risk <= cutoff)
sum(selectedLoans$profit)
table(selectedLoans$not.fully.paid)
```
We have now seen how analytics can be used to select a subset of the high-interest loans that were paid back at only a slightly lower rate than average, resulting in a significant increase in the profit from our investor's $100 investment. Although the logistic regression models developed in this problem did not have large AUC values, we see that they still provided the edge needed to improve the profitability of an investment portfolio.  

We conclude with a note of warning. Throughout this analysis we assume that the loans we invest in will perform in the same way as the loans we used to train our model, even though our training set covers a relatively short period of time. If there is an economic shock like a large financial downturn, default rates might be significantly higher than those observed in the training set and we might end up losing money instead of profiting. Investors must pay careful attention to such risk when making investment decisions.  



